\chapter{Education of Children}

We desire to tell you of some thing, 
which is of the utmost importance. We 
find it necessary, unfortunately, to repeat in a 
measure what has been told you several times. 
We speak plainly, without a flourish of words, 
because we feel our responsibility before God -- 
if we be misunderstood. We desire to remind 
you of our parish or church-school. To learn 
to read and write you send your children to 
school. You know that you must do it. But 
how many of you think of the serious obligation 
of rightly and thoroughly preparing your 
children for the life which they must live after 
only a few years? Some, indeed, give their 
attention to what they call a decent education 
for their children, for which and for whom they 
would not fall back of any one, but be as good 
and as nice as other people in town. If you 
send your children to school to study grammar 
and arithmetic, (the future mainstay of the 
``home'' are often compelled to leave their homes 
to learn even cooking and dancing), why will 
you not be just as eager to send them to school 
where they will study religion? If you are 
truly interested in the welfare of your children, 
why do you not watch as strictly, but once a week, 
how they attend to their lessons in the study of 
the Law of God, as you do in some homework, 
which the children seemed to be forced to have 
prepared within the next twelve hours for their 
public school? You must obey God, above the 
public and all other masters, or lose your souls 
for the responsibility which rests upon you for 
the present and future welfare of your children. 
Where there is intellect, there always will be 
knowledge. Still, you must educate the child. 
Teach the boy and girl geography and history; 
but if you do not train the child's will, in order 
not only to please you, its parents, but to bend 
before the holy will of Him, who is the only 
just rewarder of good and evil, then you are a 
failure as a Christian. Where there is no discipline,
there is no constancy. Where there is 
no law, there is no order, no peace, no everlasting
happiness. If no tender sympathies re-echo 
in the heart of the young, away have been cast 
the time and labor in teaching -- be it botany or 
music.

What a pity! We see young children at the 
age of ten, whose very brains seem to be rattling 
with numerical problems, while they have not 
the good manners to step out of the way of an 
old person, or even the common human feeling 
of a desire to aid in distress. I have seen even 
young men and young women stand gazing on 
one of their company, who was fainting from 
exhaustion, without the offer of the most simple 
service -- to fetch a cup of cold water. 

However regular athletic exercises are attended
to, no matter how carefully the lessons 
in physiology are prepared, little indeed, will 
they profit your children, if they know not the 
steps, up which they must climb to seek the 
Highest. If you, fathers and mothers, are 
Christians, then we ministers of the Word may 
rest in the quiet hope that your children have 
been taught dutifully and rightly to praise the 
All-majestic Creator at morn, likewise in midday, 
confessing each themselves before God, and 
openly before all men, confessing God, while at 
night they humbly implore His mercy. But if 
your children do not invoke their Guardian
Angel, if they do not bless the most pure 
Mother of our Lord Jesus Christ, and fall 
down in humble devotion, supplicating for the 
grace of our Heavenly Father, then you are not 
Christians. 

It is a sad fact, which must be recorded in 
this earthly judgment seat of God, and the 
truth of this fact is as bitter for me, as it is for 
you. Nevertheless, we acknowledge and accept 
the truthful bitterness with the hope, that it 
will prove to be a healing remedy, which will 
bring peaceful and sweet results. 

There are parents belonging to our congregation
in San Francisco, who go to the matinee 
with their children, not giving a thought before 
hand to the character of the play; they teach 
the little ladies and gentlemen, i.e. the future 
men and women of a Christian land, to buy and 
select wearing apparel, which is pleasing to the 
eyes of the world, whether it be healthful and 
sensible, or not. Yes, they are ``up to the 
times,'' they visit the classes of the public 
schools; they receive and fix their signatures 
to monthly school reports. Ah! if they would 
but fix the character of the school itself. If 
their children are tardy for five minutes they may 
not go to school without a written excuse. Yet 
it is in the power of the citizens of this country 
to have laws enacted, which would protect their 
young from being crammed with ``ologies'' and
``isms,'' and insure their healthful growth and 
the teaching of good sense. 

How is it with our church schools? All our 
children do not attend, and their friends and 
Christian neighbors do not take interest enough to 
invite them to go with their children. The parents
do not visit our school, but once a year, and 
then -- when we are not at work. What is our 
homework for the children? Only a little of 
that which is the greatest. A very little, once or 
twice a week, of the commandments of God and 
the gospel of his Son, Jesus Christ. And no 
one to think about it; no one at home to see 
that the life-work of the family is done! During
the short hour that we manage to collect a 
few unruly children, we must study and repeat, 
for even prayers are not learned in the home. 

Beloved Christians! we need your cooperation.
We may sow the seed, but -- remember -- 
the influence of your home is the sunshine 
which heats the ground. So then ask yourselves,
is not the wind too chilly, and the sun 
too low to strike its rays direct? We may trim 
the plant, but it is your duty to keep watering it. 
Oh! if you would but water the precious plants 
of your gardens with prayerful tears! We 
invite you to visit our school, from time to time,
during lesson hours. If we were asked, how 
many of us pray together with children, the 
conscientous would answer, a very, very few; 
only several in a congregation of three hundred 
souls. Generally, of an evening, the children 
are sent to bed; and sometimes some one calls 
out, say your prayers first. And from time to 
time there is a prayer, but more often there is 
only the ``saying.'' Must I explain that Christian 
children should be followed to their night's 
rest; in most cases they should be ``put'' into 
bed. 

It is the duty especially of parents to see that 
their children pray correctly, and also to pray 
with them in an audible voice themselves. Let 
this not be an act of routine. Do not for a 
moment think that it will become a daily 
routine. This reasonable discipline, when 
you kneel by the side of tender childhood 
and see the little ones pray, will lighten 
in your own heart -- at the same time 
that it does in theirs -- the fire of heavenly love. 
Moreover, your prayers must be the prayers of 
the Orthodox Church of Christ. Our Mother 
church has but one infallible model of prayer -- 
given to her by our Lord Jesus Christ. It is an 
invocation, petitions, and a doxology; in other
words, a call, a request, and a praise. If you 
will concentrate your minds upon the subject 
of each one of these divisions, then your prayer 
will not be a ``saying,'' but an ``offering.'' 
Again we ask you to give us your dutiful attention
and assistance in the work of teaching our 
children. If not for the sake of your own comfort
in old age and sickness, let us for the sake 
of their Almighty Father in Heaven, and our 
Judge, awaken in their hearts the love for that 
which is holy and truly beautiful. Amen.
