\chapter{Thoughts for Good Friday at the Passions and Burial of Christ.}

IT was our earnest desire to pursue the story 
of our Saviour's trials and His crucifixion; 
but when I looked on Him whom they pierced, 
my spirit failed before the terrible sight; I 
could not watch with Him another hour, and yet 
I could not leave the hallowed scene. It seemed 
as though I saw Him brought back from Herod 
where the soldiers mocked Him. I followed 
Him through the streets again as the cruel 
priests pushed through the wild crowd and 
hastened Him back to Pilate's court. My ears 
sounded with the cry: Crucify Him, crucify 
Him! Give us Barabbas, the robber; let Barabbas
go; but Christ, the King of the Jews, Jesus, 
the Saviour, He must die I And there He stood, 
who loved me and gave Himself for me, like a 
lamb in the midst of wolves, with none to pity 
and none to help Him. 

As Jesus Christ hung apparently helpless 
upon the cross, He had only to utter the word,
and in a moment more than twelve legions of 
angels (what an invincible force of energetic 
beings!) would be ready to succor and defend. 
But to have shunned all pain and anguish, to 
have refused the cup which His Father had 
given Him, to have rejected the cross — this 
would have been to leave man to his doom; this 
He could not do. And so, He saved others. 
Himself He could not save. 

Our Lord for a long time bore His cross, as 
though He felt not its weight; even from the 
time of the most helpless age of humanity, when 
He was born in the smallest town of the smallest 
kingdom on the earth, when there was no home, 
no cradle for Him, and when, except His humble 
mother with her guardian, none but a few shepherds
took any interest in His birth, He bore 
His cross till at length He was completely delivered
up on it. We could not follow Jesus 
throughout in His earthly life. During these 
holy days we have but endeavored to follow Him 
only through a few of the last scenes of His 
entire sacrifice. In a measure we understand, 
and we feel in His sufferings, as His body, in its 
weight drooped, being sustained but by four 
nails, as the cross was raised over the multitude 
of people on the hill and then the shock as it
went down into its socket Only a chosen few, 
and likewise in a small measure can they understand
how He — who prayed, Father forgive 
them, for they knoiv not what they do — has 
stretched out his arms on the wood in order to 
embrace a sinful world. But no mortal knoweth 
how the Word teas with God, and the Word 
was God. The Word of God is not bound by 
death. As a word from the lips dies not entirely 
away at the moment its sound ceases, but rather 
gathers new strength, and passing through the 
senses penetrates the minds and hearts of the 
hearers, so also the Hypostatical Word of God, 
the Son of God, in His saving incarnation, 
whilst dying in the flesh, fills all things with 
His spirit and might. Thus when Christ waxeth 
faint and becometh silent on the cross, then is it 
that heaven and earth raise their voice unto Him, 
and the dead preach the resurrection of the 
crucified, and the very stones cry out. And the 
sun was darkened, and the veil of the temple 
was rent in the midst; and the earth did quake 
and the rocks rent; and the graves were opened, 
and many bodies of the saints which slept arose. 
O sinful man, O nature, bereft of perfection. 
O reason, a mind earthly winged, down low, 
stoop thou under cover of repentant shame before
the light of this grave. Christian, there is 
no other place for thee to-day but by the Cross 
of Christ. " Broken and distributed is the Lamb 
of God, which is broken, yet not severed, which 
is ever eaten, yet never consumed, but sanctifying
those that participate." Therefore come, ye 
sons of toil and. ye daughters of Eve; come, 
come, ye citizens of the easy-going wide world; 
see, His side is now and forever opened for us. 
O mother, sufferest thou for thy children? Bring 
them to the tomb of Jesus and quiet thyself in 
the stillness of His silence. Brother, sister, 
behold in the Saviour thy kinsman. And thou 
poor, lonely wanderer, here He quietly lies in 
one place that thou may est find Jesus, thy only 
Friend. We come, who labor in Thy infinite 
sorrow for the sins of mankind, we who are 
heavy laden with our infirmities, we come and 
supplicate before Thy breathless form. 

Uniting all things in one, grant that we all 
may inseparably be one with Thee and Thy 
Father, O Lord Jesus Christ! Thou that reconcilest
all, grant that all may be of one mind in 
faith and in love toward Thee. Thou that bears 
not with the envious and contentious, destroy all 
wicked heresy and separations. O Jesus! Thou 
that lovest and pitiest, gather into one flock all
wandering sheep. Thou that givest peace to all, 
still the voice of spite and dispute among those 
who call upon Thy name. Thou who communicatest
to us the very Body and very Blood of 
Thyself, grant that we truly be flesh of Thy 
flesh and bone of Thy bones. O Jesus, the God 
of our hearts, unite us with Thee, now and forever.
Amen.
