\chapter{The Condition of Society.}

HOW long will it thus go on! When will 
the baptized become active Christians, so 
that the pastors may give their attention to the 
conversion of the heathen? What a terrible 
battle we must fight. Already the fire of hell 
is in the world. Great cities are multiplying 
throughout the land. The farmer, as the word 
is defined in our dictionaries, is a thing of the 
past. It is now the land-owner with a mansion 
in the city, a yacht on the sea, and with a private 
train across the continent. There are comparatively
but a few laborers in the fields — too poor 
to support families. The quiet country homes 
are becoming few, shall I say precious? I fear 
not so, because people are fast losing their 
ability to rightly estimate the value of things. 
Most of the cities in all the world are overcrowded.
The female portion of the population 
is most conspicuous. A stupid craze after 
unwholesome fashions is the one all-absorbing 
passion of the majority of women. There is no 
room for gardens and yards ; most of the children
in San Francisco are actually brought up 
in the streets. Oh, how few of them feel the 
blessed influence of a Christian home! Young 
men and young women are continually" on the 
go," as they say. And this "go" is a nervous, 
unsteady rush to" keep up with the times." And 
after all their hurry nothing is left but steam 
and vapor, for they are empty, as empty as the 
changing and vanishing world can be. Yet they 
fret and inquire: "Where shall we go to and 
what shall we see? What shall we do? Oh! 
what can we do?" If you promenade along the 
broad avenue or pass through the narrow lane, 
if you visit the meeting halls in the city or look 
into the factories, everywhere you see that same 
all-devouring gaze of the bold young woman, 
who stares with a kind of artificial movement of 
the eyes. And sometimes you hear even so-called
Christians say that it is a weakness of 
character in one who has the downcast eyes of 
modesty, the blush of innocence. Such people 
do not know the live sense and fine impulse of a 
pure conscience. When a young man puffs 
tobacco smoke or shows his teeth with a disapproving
smile in the presence of and at the 
conversation of older people, then society is
wrong; something is the matter with his family. 

In view of all this, beloved, the preacher of 
the Word of God is obliged by a terrible oath 
he has given before he received the gift in 
Apostolic succession at his ordination, to present 
to you the whole of the Truth, not a part of it. 

The number of unmarried people is increasing.
And there are some married people who 
say: "We do not want children, because we 
want to have as much pleasure as possible." 
This is a false position, for in a Christian marriage
one kind of pleasure is not allowed continually.
Christians marry for the sake of God 
and His law as much as they do for themselves. 
But Christians who remain single renounce 
marriage and live holy for the sake of God and 
Him alone. Thus we find that the family tie is 
abused, as well as the single state. Courtship 
of young people just out of school is not to be 
advised, because it often leads to debauchery. 
A courtship running through long years also 
gives occasion to sin and a species of wrongdoing
to God, for the heart and its love are 
stolen from God and thrown away on a man. 

Throughout all the long centuries of Christianity
there have been in the Church heroic 
members, young people of both sexes, who by
the grace of God have kept their souls pure and 
intact, and have dedicated to the honor of God 
the noblest attribute of their human life, namely, 
an untarnished purity of soul and body. Such 
persons have had the courage and such unbounded
confidence in God's assistance that, 
although living in the world and its dangers, 
though threatened by the cravings of their own 
individual passions and by the temptations of 
the devil, yet they have succeeded bravely in 
preserving this treasure even in a frail earthen 
vessel, have carried it uninjured through life's 
long journey here below, and have finally presented
it to their Lord. 

Christian heroes and heroines, you who have 
imitated or who still do imitate the sublime example
of the Most Blessed Virgin, the Church 
admires your spirit of sacrifice as she does that 
of the holy martyrs, who in a few hours finished 
their contest and proved their fidelity to God 
and their faith; because you have to combat, to 
suffer, and to sacrifice your whole life through. 
With joy and veneration do the angels look down 
upon you, for you resemble themselves. With 
motherly affection and with mighty power does 
the Holy Virgin Mary when you earnestly pray 
throw her sheltering omophorion around you, 
for you are her pupils and imitators. "With the 
sweetness of divine love the heavenly Bridegroom 
will fill your heart and more than compensate 
you for the fleeting, transient, worldly love that 
you have laid down at His feet. The eternal 
Judge will find you waiting like the wise and 
prudent virgins who all through life carry in 
their hands the pure oblation of love and the 
burning light of good example. Therefore, faithful
to the end, He will invite you to the eternal 
wedding feast in heaven. Amen. 
