\chapter{Address on Christmas Day.}

THE Orthodox Catholic Church celebrates 
to-day the nativity of our Lord and Savior 
Jesus Christ. 

Christmas day is generally observed as a holiday
throughout the world. And verily it should 
be a day of rejoicing. I say rejoice, for, glory 
to God in the highest, we are now enabled to 
obtain peace on earth, as the good will of our 
Creator and Judge has been proclaimed to all 
men. 

Unfortunately we now often see in the world 
a decline in religious celebrations of Christian 
festivals,\footnote{The majority of churches in San Francisco (as is the 
general rule in America) are closed and have no purely 
religious services whatever on the 25th of December.}
just as in political and civil holidays 
we now often notice a tendency to withdraw from 
public scenes to one's own narrow sphere of 
privacy, personal comfort, and individual satisfaction.
Such conditions are detrimental to society
at large. The first cause of it all is the 
family — the home from which all men and 
women take into the world the form of their 
future character. The nation which is devoid 
of lessons in religion and patriotism in the 
family — which is the fountain source of all 
learning — such a nation is going to destruction. 
Since we to-day magnify Christ, the giver of 
life, who now was born in the flesh for our sake 
of the unwedded and most pure Virgin Mary, 
we may most appropriately, as duty bound, consider
the worth of our celebration. When we 
hear the teaching of Him who to-day was a little 
child seeking shelter with the cattle and sheep, 
and remember that He said, I am the way, and 
the life, and the truth, and also, the truth will 
make you free, we are impressed with the exalted 
character of Him whose mission it was and is 
to gather all mankind into the one fold of the 
Only Shepherd. And our Lord says the truth 
will save us. We must study the truth to the 
full measure of our competency, and know our 
Lord Jesus Christ the Savior. We must use 
our God-given intellect and discern true doctrine 
from falsehood. We must worship God with our 
whole being, body and soul, consequently we 
develop our reason (as our faith grows stronger) 
and praise the Lord understandingly. For when 
we worship with the inspiring feeling of our 
heart only, we bring a half offering. The offering
must be whole. 

When we study the different forms of worship 
of different people who are considered Christians, 
we see that some bow in adoration with fervent 
devotion, while others with exultant heart sing 
loud praise, as if carried away in ecstasy, depriving
the spirit of one of its faculties, i.e., 
adoration. Now, this is according to science, 
and those who have studied psychology know it. 
True and complete science is always compatible 
with religion. The Orthodox and Apostolic 
Church has ever held that man must offer praise 
from a warm heart while devoutly adoring the 
Sublime both physically and spiritually, and at 
the same time following truthful doctrine. Yes, 
we must seek and study the truth. I do not say 
we can at once perfectly know it, for we must 
first comprehend God, who is the truth. When 
we become Christ-like we know the truth, and 
the truth makes us free. 

The Church is the treasure house of God upon 
earth. I offer these precious gems to you, hoping 
you will take them away to your homes, safely 
guarded in a grateful heart, and earnestly ponder 
over them, seeking God in prayer accompanied 
by a pure life. For only then we may enter
into the true spirit of this celebration and rightly 
comprehend the hymn we sing : " Christ is born ! 
Ye faithful, glorify! Christ from the heavens, 
Oh come to greet Him ! Christ upon the earth, 
be ye lifted on high! Sing to the Lord, all the 
earth ! And in gladness praise him. O ye nations!
for he hath been glorified!" Amen.

